---
title: "Engaging and Beautiful Data Visualizations with ggplot2"
subtitle: "Fundamentals & Workflows: Exercises"
author: "Your Name"
date: last-modified
format: 
  html:
    toc: true
execute:
  message: false
  warning: false
  fig-width: 10
  fig-height: 6.2
editor: source
---

# Exercise 1

-   **Discuss / investigate with your neighbor:**
    -  What are the differences between `geom_line()` and `geom_path()`?
    -  Why can you use `geom_smooth()` and `stat_smooth()` interchangeably?
    -  What are the three ways to *remove a legend* from a ggplot?
    -  Which is your favorite theme? 
    -  What is your favorite ggplot2 extension package?


*Add your answers here.*


# Exercise 2

-   **Explore the TfL bike share data visually:**
    -  Create a *time series* of counts per day and night.
    -  Draw *box and whisker plots* of average temperatures per month.
    -  Visualize bike counts per weather type and period as *bar chart*. 
-   **Combine the three plots with patchwork.**
-   **Export the final graphic in a format of your choice.**


## Data import

```{r}
#| label: preparation
library(readr)
library(ggplot2)

bikes <-
  read_csv(
    here::here("data", "london-bikes.csv"),
    col_types = "Dcfffilllddddc"
  )
```


## Visualizations

### Time series of counts per day and night

```{r}
#| label: timeseries

```

### Box and whisker plots of average temperatures per season

```{r}
#| label: boxplot

```

### Bar chart of counts per weather type and period of the day

```{r}
#| label: barchart

```


## Combine Plots

```{r}
#| label: combined-graphic
library(patchwork)

```


## Save Multipanel-Plot

```{r}
#| label: export-graphic

```
