---
title: "Engaging and Beautiful Data Visualizations with ggplot2"
subtitle: "Working with Colors: Exercises"
author: "Your Name"
date: last-modified
format: 
  html:
    toc: true
execute:
  message: false
  warning: false
  fig-width: 7.5
  fig-height: 4.7
editor: source
---

```{r}
#| label: preparation
library(readr)
library(dplyr)
library(ggplot2)

## data
bikes <-
  read_csv(
    here::here("data", "london-bikes.csv"),
    col_types = "Dcfffilllddddc"
  )
```


# Exercise 1

-   **Add colors to our bar chart from the last exercise:**
   -   Encode the bars and direct labels by color for "fair weather" (clear, scattered clouds, broken clouds) and "grey weather" (rain, cloudy, snow)
   -   Use the same colors to encode the respective text bits in the title ("Fair weather preferred") and subtitle ("rainy, cloudy, or snowy days").


## Previous Bar Chart

```{r}
#| label: bars-weather-2015
bikes |> 
  filter(year == "2015") |> 
  mutate(weather_type = forcats::fct_reorder(weather_type, count, .fun = sum)) |> 
  ggplot(aes(x = count, y = weather_type)) +
  stat_summary(
    geom = "bar", fun = sum, 
    color = "grey20", fill = "beige", width = .7
  ) +
  stat_summary(
    geom = "text", fun = sum,
    aes(
      label = after_stat(paste0("  ", sprintf("%2.2f", x / 10^6), "M  ")),
      hjust = after_stat(x) > .5*10^6
    ),
    family = "Asap SemiCondensed"
  ) +
  scale_x_continuous(
    expand = c(0, 0), name = "**Reported bike shares**, in millions", 
    breaks = 0:4*10^6, labels = c("0", paste0(1:4, "M"))
  ) +
  scale_y_discrete(labels = stringr::str_to_sentence, name = NULL) +
  coord_cartesian(clip = "off") +
  labs(
    title = "Fair weather preferred—even in London",
    subtitle = "Less than 10% of TfL bikes shares have been reported on rainy, cloudy, or snowy days in 2015.",
    caption = "**Data:** Transport for London / freemeteo.com"
  ) +
  theme_minimal(base_size = 14, base_family = "Asap SemiCondensed") +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank(),
    axis.title.x = ggtext::element_markdown(hjust = 0, face = "italic"),
    axis.text.y = element_text(color = "black", size = rel(1.2)),
    plot.title = element_text(face = "bold"),
    plot.subtitle = element_text(margin = margin(b = 20)),
    plot.title.position = "plot",
    plot.caption = ggtext::element_markdown(color = "grey40")
  )
```



# Exercise 2

-   **Create a plot of your choice with a sequential (non-default) color palette.**
    -   Inspect the HCL spectrum. Adjust the palette if needed.
    -   Test the palette with regard to colorblindness. Adjust the palette if needed.
    -   Save and share the graphic.


## Visualization

```{r}
#| label: plot-colors

```

## Evaluate HCL Spectrum

```{r}
#| label: hcl-spectrum

```

## Emulate CVD

```{r}
#| label: hcl-spectrum

```

## Save Graphic

```{r}
#| label: export-plot

```
